---
title: "Individual Assignment"
author: "Hamdha Saheer"
date: "2026-02-05"
output: html_document
---


   Financial Valuation and Analysis Using R 
   
   
SESSION 01: BOND VALUATION SCENARIO   
     
     Sunrise Holding PLC
     
```{r}
# Bond Parameters
face_value <- 5000      # Face value of the bond
coupon_rate <- 0.07      # Annual coupon rate (7%)
market_rate <- 0.09       # Market interest rate (9%)
years_to_maturity <- 10   #  maturity of 10 years

# Calculate annual coupon payment
coupon_payment <- face_value * coupon_rate

# calculates the present value of all future coupon payments from a bond
pv_coupons <- sum(coupon_payment / (1 + market_rate)^(1:years_to_maturity))

# Calculate present value of face value
pv_principal <- face_value / (1 + market_rate)^years_to_maturity

# Total fair value of bond
fair_value <- pv_coupons + pv_principal

# Output the result
cat("Annual Coupon Payment: LKR", coupon_payment, "\n")
cat("PV of Coupon Stream:   LKR", round(pv_coupons, 2), "\n")
cat("PV of Principal:       LKR", round(pv_principal, 2), "\n")
cat("Fair Value of the Bond: ", round(fair_value, 2), "\n")

```
When the Fair value of the bond (which is LKR 4358.23) is less than Face value of the bond (which is LKR 5000, at this instant, the bond is said to trade at a “Discount”. This status arises because the market yield currently prevailing at 9% , it is higher than the annual coupon payments calculated at a rate of 7%. If market interest rates rise further, the value of this bond will decrease. Then market rates will fall, and the bond’s price will increase.







SESSION 02: EQUITY VALUATION SCENARIO

    Oceanic Beverages PLC
    
```{r}
# Scenario 03: Non-Constant growth in dividends
# Inputs
D0 <- 3.5           # Last dividend paid
g1 <- 0.12         # Growth rate for first 4 years
g2 <- 0.05         # Growth rate thereafter
r  <- 0.14         # Required rate of return

# Step 1: Calculate dividends for years 1 to 4
D1 <- D0 * (1 + g1)
D2 <- D1 * (1 + g1)
D3 <- D2 * (1 + g1)
D4 <- D3 * (1 + g1)

# Step 2: Calculate terminal value at end of year 4
TV <- D4 * (1 + g2) / (r - g2)

# Step 3: Discount dividends and terminal value to present
PV_D1 <- D1 / (1 + r)^1
PV_D2 <- D2 / (1 + r)^2
PV_D3 <- D3 / (1 + r)^3
PV_D4 <- D4 / (1 + r)^4
PV_TV <- TV / (1 + r)^4

# Step 4: Calculate total present value (share price)
fair_value <- PV_D1 + PV_D2 + PV_D3 + PV_D4+ PV_TV

# Output result
cat("Projected D1-D4:", round(D1,2), round(D2,2), round(D3,2), round(D4,2), "\n")
cat("Terminal Value at Year 4: LKR", round(TV, 2), "\n")
cat("Fair Value per Share: LKR", round(fair_value, 2), "\n")
```
	The valuation is highly sensitive to the perpetual growth rate and the required rate of return. Because the terminal value constitutes most of the share price, a small increase in the required return or a decrease in long-term growth would lead to a substantial drop in the fair value per share.



SESSION 03: ANALYSIS OF HISTORICAL INVESTMENT PERFORMANCE

     Tech Nova PLC
    
```{r}
# Annual returns in decimal form
returns <- c(0.18, 0.10, 0.04, 0.12, 0.20, 0.08)

# Calculate average return
average_return <- mean(returns)

# Calculate sample standard deviation (risk)
std_deviation <- sd(returns)

# Display results in percentage format
cat("Average Return:", round(average_return * 100, 2), "%\n")
cat("Standard Deviation (Risk):", round(std_deviation * 100, 2), "%\n")
```

•	These measures quantify the trade-off between reward and uncertainty. A 12% mean return coupled with a 6.07% standard deviation indicates moderate volatility; The actual return has historically deviated from the average by roughly 6% in either direction.




       Eco Grow PLC
  Projected Return Analysis under Uncertain Outcomes
```{r}
# Define returns and probabilities
returns_f <- c(0.25, 0.12, 0.05)         # Convert % to decimals
probabilities <- c(0.25, 0.50, 0.25)   # Probabilities must sum to 1

# Step 1: Expected return
expected_return <- sum(returns_f * probabilities)

# Step 2: Variance
variance <- sum(probabilities * (returns_f - expected_return)^2)

# Step 3: Standard deviation
std_deviation <- sqrt(variance)

# Output results
cat("Expected Return:", round(expected_return * 100, 4), "%\n")
cat("Variance:       ", round(variance * 10000, 2), "\n")
cat("Standard Deviation (Risk):", round(std_deviation * 100, 4), "%\n")
```

	The distribution of returns influences the perceived risk. While Eco Grow PLC offers a higher expected return(13.5%) than Tech Nova PLC (12%), it also presents higher risk as evidences by its standard deviation (7.23% vs 6.07%). The 25% probability of achieving only a 5% return highlights the potential for significant downside compared to the expected mean.


